Bootstrapping Multi-atlas Hippocampal Segmentation with \mb

### 2013-04-13 paper.Rnw
Fixed results section order, and exported some analysis code to files

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable}
\usepackage{titlesec, siunitx}
\usepackage[round,authoryear]{natbib}

@@

In this experiment, the five $300\mu$-isotropic voxel Winterburn atlases are
used as the atlas library for \mb segmentation. LOOCV is conducted separately
for two different input datasets: the four Winterburn atlas subject BRAVO MR
images (referred to as the {\em Winterburn BRAVO} dataset), and the five
Winterburn atlas subject images resampled to 0.9mm-isotropic voxel resolution
(referred to as the {\em Winterburn Resampled} dataset. Each subject in the
dataset is segmented by \mb with that subject's image excluded from the atlas
library. The template library consists of 3T T1 images (0.9mm-isotropic
voxels) of healthy subjects in addition to the images from the dataset being
evaluated.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                               Results
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

%%%%%
%%%%%                  Results: ADNI-1 cross-validationn                   %%%%%
%%%%%
\subsection{ADNI1 Cross-Validation}

%                                                                              %
% ADNI1 Complete 1Yr Validation - results preparation                             %
%                                                                              %
<<ADNI1-xval-load, include=FALSE, cache=TRUE>>=
# all data from the validation trials
a_data       = read.csv(gzfile('data/a2a_ants/results_2013_01_03.csv.gz'))
#t_data       = read.csv(gzfile('data/a2a_tracc/xfmtrials_results_2013-04-01.csv.gz'))
t_data       = read.csv(gzfile('data/a2a_tracc/results_xfmjoin_2013_04_04.csv.gz'))
diagnoses    = read.csv('data/a2a_diagnoses.csv')
@

<<ADNI1-xval-prep,dependson='ADNI1-xval-load',include=FALSE,cache=TRUE>>=
# simplify the data
t_data       = subset(t_data, atlases %in% c(1,3,5,7,9))
a_data       = subset(a_data, atlases %in% c(1,3,5,7,9))
t_data$se    = NULL
t_data$sn    = NULL
t_data$j     = NULL
a_data$se    = NULL
a_data$sn    = NULL
a_data$j     = NULL

# add registration method column 
t_data       = cbind(t_data, reg_method = "ANIMAL")
a_data       = cbind(a_data, reg_method = "ANTS")

# separate MA majvote from MB so we can compute differences
# we only compare multiatlas using majority vote
t_ma = subset(t_data, approach=="ma" & method=="majvote")
t_ma$approach="multiatlas"
a_ma = subset(a_data, approach=="multiatlas" & method=="majvote")
t_mb = subset(t_data, approach=="mb")
a_mb = subset(a_data, approach=="mb")

ma = rbind(t_ma,a_ma)
mb = rbind(t_mb,a_mb)

# set up a few equivalences for multi-atlas, to make nomenclature simpler
ma$templates =  ma$atlases  # because no template library
mb[mb$method != 'majvote',]$templates = mb[mb$method != 'majvote',]$top_n

# aggregate over batches
ma_mean = aggregate(k ~ subject + reg_method + approach + method + atlases + templates + top_n + label, data = ma, mean)
mb_mean = aggregate(k ~ subject + reg_method + approach + method + atlases + templates + top_n + label, data = mb, mean)

all_data = merge(ma, mb, by=c("timestamp", "atlases", "batch", "label", "reg_method", "subject"), 
                 suffixes=c(".ma", ".mb"))

all_data_mean = merge(ma_mean, mb_mean, by=c("reg_method", "label", "atlases", "subject"), 
                 suffixes=c(".ma", ".mb"))

all_data$k_diff = all_data$k.mb - all_data$k.ma
all_data_mean$k_diff = all_data_mean$k.mb - all_data_mean$k.ma

vote_levels <- list("Majority Vote" = "majvote", 
                      "NMI Vote" = "nmi", 
                      "Cross-correlation Vote" = "xcorr")
levels(all_data$method.mb) <- vote_levels
levels(all_data_mean$method.mb) <- vote_levels

all_data       = merge(all_data, diagnoses)
all_data_mean  = merge(all_data_mean, diagnoses)
remove(t_ma, a_ma, t_mb, a_mb, ma, mb)
@
%                                                                              %
%                                                                              %

%
% synonyms for significant: 
% powerful, suggestive, apparent, definite, evident, large, marked, material, 
% noticeable, plain, pronounced, substantial, substantive, 
%
% antonyms: 
% negligible, minor, no appreciable, scant

@@

\begin{figure}
<<ADN1-xval-k-mean, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
ggplot(all_data_mean, 
       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + 
  facet_grid(reg_method~method.mb) + 
  stat_summary(fun.y=mean,geom='line',
               aes(y=k.mb, weight=1)) + 
  #stat_smooth(method='lm', formula=y~log(x)) + 
  scale_colour_hue(name="Atlases") +
  xlab( "Number of Templates" ) + 
  ylab( "Mean DSC" ) + 
  opts(legend.direction = "horizontal", legend.position = "bottom")
@
  \label{fig:ADNI1-xval-k_mean}
  \caption{
  {\bf Mean DSC by atlas and template library size, registration algorithm,
  and label fusion method.} The mean DSC score is calculated over all subjects 
  from the ADNI1 Cross-Validation dataset for each parameter setting.}
\end{figure}

@@

\begin{figure}
<<ADN1-Xval-k-diff, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
ggplot(subset(all_data_mean, reg_method=="ANTS" & method.mb=="Majority Vote" &
                templates.mb %in% seq(1,20,2)), 
       aes(x=templates.mb, y=k_diff, colour=as.factor(atlases))) + 
  #facet_grid(reg_method~method.y) + 
  stat_summary(fun.data=mean_cl_boot,geom='errorbar',
               aes(y=k_diff,colour=as.factor(atlases), width=0.2)) +
  stat_summary(fun.y=mean, geom="point", 
               aes(y=k_diff,colour=as.factor(atlases))) + 
  stat_summary(fun.y=mean, geom="line", 
               aes(y=k_diff,colour=as.factor(atlases))) + 
  #stat_smooth(method=lm, formula=y~log(x)) + 
  geom_hline(aes(alpha=0.5, yintercept=0), linetype='dashed') + 
  scale_x_continuous(breaks=seq(1,20,2)) + scale_y_continuous(breaks=seq(-1,1,by=0.01)) + 
  scale_colour_hue(name="Atlases") +
  scale_alpha_continuous(legend=FALSE) + 
  scale_linetype_discrete(legend=FALSE) + 
  xlab( "Number of Templates" ) + 
  ylab( "Increase in mean DSC between MAGeT brain and multiatlas" ) +
  opts(legend.direction = "horizontal", legend.position = "bottom")
@
  \caption{The difference in mean DSC between \mb and multi-atlas segmentations for 
  a range of parameter settings.}
  \label{ADNI1-xval-k-diff}
\end{figure}

@@

\begin{figure}
<<ADN1-Xval-variability, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
ANTSmajvote = subset(all_data, 
                     reg_method=="ANTS" & 
                     method.mb=="Majority Vote" &
                    templates.mb %in% seq(1,20,2))

@@

ggplot(stats.tests, 
  aes(x=templates.mb, y=p.value, colour=as.factor(atlases), 
      size = abs(statistic))) + 
  facet_grid( . ~ stat) + 
  geom_line(aes(size=0.1,area=0.1)) + 
  geom_point(aes(shape = factor(direction))) + 
  geom_hline(aes(yintercept=0.05, alpha=0.5), linetype='dashed') + 
  geom_hline(aes(yintercept=0.01, alpha=0.5), linetype='dashed') + 
  scale_shape_discrete("t-statistic sign", labels=c("positive","negative")) + 
  scale_size_area(name="t-statistic magnitude") + 
  scale_x_continuous(breaks=seq(1,20,by=2)) + 
  scale_y_log10(limits=c(0.001,1), breaks=c(0.01,0.05,0.1,1.0)) + 
  xlab( "Number of Templates" ) + 
  ylab( "p-value" ) +
  scale_colour_hue(name="Atlases") +
  scale_alpha_continuous(guide='none') + 
  scale_linetype_discrete(guide='none') + 
  scale_size_continuous(guide='none') + 
  opts(legend.direction = "horizontal", legend.position = "bottom")
@
  \caption{{\bf Difference in variability of \mb vs. multi-atlas segmentation 
  accuracy.}
  Variability of segmentation accuracy within 10 rounds of validation per subject. 
  Standard deviation and variance is computed per subject for both \mb and 
  multi-atlas, and the distribution of these test statistics is compared with a 
  t-test. The p-value of this test, is shown on the y-axis (scaled logarithmically).}
  \label{ADNI1-xval-variability}
\end{figure}

% <<ADNI1-Xval-sd-by-subject, cache=TRUE>>=
% SDs = data.stats[c('subject','label','atlases','templates.mb','MA.sd','MB.sd')]
% SDs = subset(SDs, templates.mb %in% c(5,10,15,20) & atlases == 5)
% 
% melted = melt(SDs, measure.vars=c('MA.sd','MB.sd'))
% ggplot(melted, aes(x=templates.mb, y=value, 
%                    group=interaction(templates.mb, variable), colour=as.factor(variable))) + 
%   scale_x_continuous(breaks=c(5,10,15,20)) + 
%   geom_violin() 
% @
% \begin{figure}
% <<ADN1-Xval-k-diff-ANTS-20tmpls, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
% a_20_tmpls = subset(all_data_mean, 
%                       reg_method  == "ANTS" &
%                       method.mb   == "Majority Vote" &
%                       templates.mb== 20)
% ggplot(a_20_tmpls, aes(x=factor(atlases))) +
%   #geom_point(aes(y=k.mb,group=1,colour="MAGeT"),position="jitter") + 
%   #geom_point(aes(y=k.ma,group=2,colour="MA"),position="jitter") + 
%   geom_violin(aes(y=k.mb,group=factor(atlases),colour="MAGeT")) + 
%   geom_violin(aes(y=k.ma,group=factor(atlases),colour="MA")) + 
%   geom_smooth(aes(y=k.mb,group=1,colour="MAGeT")) +
%   geom_smooth(aes(y=k.ma,group=2,colour="MA")) +
%   
%   #stat_summary(fun.y=mean,geom='point',size=3,colour="red",aes(y=k.ma)) +
%   #stat_summary(fun.y=mean,geom='line',colour="red",aes(y=k.ma,group=1)) +
%   #stat_summary(fun.y=mean,geom='point',size=3,colour="blue",aes(y=k.mb)) +
%   #stat_summary(fun.y=mean,geom='line',colour="blue",aes(y=k.mb,group=1)) +
%   labs(x="Number of Atlases", y = "Kappa", 
%        title="mean Kappa of MA and MAGeT brain (ANTS/20 templates) \nvs\natlas library size") +
%   scale_colour_hue(name="Method") + 
%   coord_cartesian(ylim=c(0.7,0.9))
%   #theme(legend.justification=c(1,0), legend.position=c(1,0))
% @
%   \caption{Change in mean Kappa between multi-atlas and \mb using \ants,
%   20 templates}
%   \label{}
% \end{figure}
% 
% \begin{figure}
% <<ADN1-Xval-ANTS-20tmpls-t-test-sd, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
% a_20_tmpls = subset(all_data, 
%                       reg_method  == "ANTS" &
%                       method.mb    == "Majority Vote" &
%                       templates.mb == 20 &
%                       atlases     == 9 & 
%                       volume.ma < 10000)
% stddev = ddply(a_20_tmpls, c("subject", "label"), function (df) {
%   data.frame(
%     MA = sd(df$k.ma),  #MA
%     MB = sd(df$k.mb)  #MB
%   )
% })
% 
% 
% # TODO: against SNT
% melted = melt(stddev, id.vars=c("subject", "label"))
% ggplot(melted, aes(x=variable, y=value)) + geom_boxplot()
% t.test(stddev$MA, stddev$MB)
% @
%   \caption{}
%   \label{}
% \end{figure}
% 
%

\begin{figure}
<<ADN1-xval-Bland-Atlman, cache=TRUE, dependson='ADNI1-Xval-prep', fig.width=8, fig.height=8>>=
df_subset = subset(all_data, 
                      reg_method   == "ANTS" &
                      method.mb    == "Majority Vote" &
                      templates.mb %in% c(1,11,19) &
                      atlases      %in% c(1,5,9) &
                      volume.ma     < 10000)

@@

ggplot(hc_vol, aes(x=ma, y = diff, group=Atlases)) + 
    geom_hline(aes(yintercept = mean + c(-2, 0, 2) * sd, colour='brown'), 
                linetype=2) + 
    geom_point() + 
    geom_smooth(method="lm") + 
  facet_grid(Atlases~Templates, labeller = label_both) + 
  xlab(expression('Mean manual and MAGeT-Brain total hippocampus volume ' (mm^3))) + 
  ylab(expression('Manual - MAGeT-Brain total hippocampus volume ' (mm^3))) 
@
  \caption{Bland-Altman plots comparing manual and \mb hippocampal volumes 
  when using a varying number of atlases and templates.}
  \label{ADNI1-xval-Bland-Altman}
\end{figure}
% 
% 
% \begin{figure}
% <<ADN1-Xval-k-by-disease, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
% a_20_tmpls = subset(all_data, 
%                       reg_method  == "ANTS" &
%                       method.y    == "Majority Vote" &
%                       templates.y == 20 &
%                       atlases %in% c(1,9))
% template_library = read.csv('data/a2a_ants/template_library_composition.csv')
% template_library$subject = NULL
% a_20_tmpls = merge(a_20_tmpls, template_library,
%                    by.x=c('timestamp', 'templates.y'), 
%                    by.y=c('timestamp', 'templates'))
% ggplot(a_20_tmpls, aes(y=k.y, x=templates_CN,colour=DX)) +  # why does k.x look the same?
%   facet_grid( . ~ atlases) + 
%   geom_smooth(method='lm',fullrange=F) 
%  # add 
% @
%   \caption{Mean kappa vs. proportion of disease category in template library by disease category}
%   \label{}
% \end{figure}
% 
% 
% \todo{show cost (in registrations) / benefit tradec off graph:  show number of
% registrations per Kappa?  or hours of manual labour per Kappa?)}
% discuss run-time for each of MAGet and multi-atlas

%%%%%
%%%%%                  Results: ADNI-1 Complete 1Yr Validation               %%%%%
%%%%%
\subsection{ADNI1 Complete 1Yr Validation}

<<ADNI-src-volumes-prep,echo=F,cache=T>>=
#bl    = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
mb    = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_MAGeT_volumes.csv")
maper = read.csv("data/ADNI_baseline_volumes/MAPER_volumes.csv")
snt   = read.csv("data/ADNI_baseline_volumes/UCSFSNTVOL.csv")
fs    = read.csv("data/ADNI_baseline_volumes/UCSFFSX_02_15_12.csv")
fsl   = read.csv("data/ADNI_baseline_volumes/ADNI1_1.5T_FSLFIRST_volumes.csv", sep=';')
qc    = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_QC_Park.csv", sep="\t")

# Get subjects, and diagnoses
yr1 = read.csv("data/ADNI_baseline_volumes/ADNI1_Complete_1Yr_1.5T_11_15_2012.csv")
yr1 = subset(yr1, !grepl("Scaled_2", Description)) #TODO: remove Scaled_2? 
yr1$VISCODE <- factor(yr1$Visit, levels = c(1,2,3,4), labels=c("bl","m03","m06","m12"))
subjects = yr1[,c("Image.Data.ID", "RID","VISCODE")]  # we only need these fields
dx = unique(adnimerge[c("DX.bl", "RID", "VISCODE")])
subjects = merge(subjects, dx)

# fetch just the columns we need, and do a little renaming
mb_pruned     = data.frame(MAGeT_L = mb$Left_Hippo, 
                           MAGeT_R = mb$Right_Hippo, Source = mb$SourceImageID)
maper_pruned  = data.frame(MAPER_L = maper$Left_Hippo, 
                           MAPER_R = maper$Right_Hippo, Source = maper$SourceImageID)
snt_pruned    = data.frame(SNT_L = snt$LEFTHIPPO, 
                           SNT_R = snt$RIGHTHIPPO, Source = snt$IMAGEUID)
fs_pruned     = data.frame(TEMPQC =fs$TEMPQC, 
                           FS_L=fs$ST29SV, 
                           FS_R=fs$ST88SV, 
                           Source = fs$IMAGEUID)
fsl_pruned    = data.frame(FSL_L=fsl$X17, 
                           FSL_R=fsl$X53, Source=fsl$Source)

# Now create one data.frame with all the yr1 volume data we have from each datasource
combined = merge(subjects, mb_pruned   , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
combined = merge(combined, maper_pruned, by.x="Image.Data.ID", by.y="Source", all.x=TRUE)  # only Baseline
combined = merge(combined, snt_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
combined = merge(combined, fs_pruned   , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
combined = merge(combined, fsl_pruned  , by.x="Image.Data.ID", by.y="Source", all.x=TRUE)
combined = merge(combined, qc          , by=c("Image.Data.ID", "RID"), all.x=TRUE)

# Step 2: Of those, get counts for the numbers that fail QC
package_totals = data.frame(
  Total     = c(length(subjects$VISCODE), 'N/A'),
  SNT        = c(length(which(!is.na(combined$SNT_L)))    , "--"),
  MAGeT      = c(length(which(!is.na(combined$MAGeT_L)))  , length(which(combined$MAGeT_QC==0))), 
  MAPER      = c(length(which(!is.na(combined$MAPER_L)))  , "--"), 
  FSL        = c(length(which(!is.na(combined$FSL_L)))    , length(which(combined$FSL_QC==0))), 
  FS         = c(length(which(!is.na(combined$FS_L)))     , length(which(combined$TEMPQC == "Fail"))),
  row.names  = c("Images", "Failures")
)

all_qc = subset(combined, TEMPQC != "Fail" & MAGeT_QC == 1 | FSL_QC == 1 )

attach(all_qc)
totals = data.frame(RID = RID, VISCODE = VISCODE, DX = factor(DX.bl),
                  MAGeT = MAGeT_L + MAGeT_R, 
                  MAPER = MAPER_L + MAPER_R, 
                  SNT   = SNT_L + SNT_R, 
                  FS    = FS_L + FS_R, 
                  FSL   = FSL_L + FSL_R)
detach(all_qc)
totals.complete = na.omit(totals)
@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1 1yr Complete Dataset Package Totals & Failures                        %
%                        
%
<<ADNI1-scr-package-totals, echo=F, results="asis",cache=FALSE>>=
caption = paste(
  "Pass/fail quality control indicators were supplied with the FreeSurfer",
  "volumes downloaded from the ADNI website (we used the temporal lobe quality",
  "control indicator, TEMPQC). One of the authors (MP) performed visual",
  "quality inspection for MAGeT and FSL segmentations.", sep=" ")
latex(package_totals, file="", size="scriptsize", 
    caption = caption,
    title = "",
    label="tab:ADNI1-1.5T-Complete-1Yr-Dataset-Method-Totals")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{algorithm, color, algpseudocode, amsmath, url, geometry, ctable}
\usepackage{titlesec, siunitx}
\usepackage[round,authoryear]{natbib}
\usepackage[section]{placeins}  % keep floats in their place.

@@

In this experiment, the five $300\mu$-isotropic voxel Winterburn atlases are
used as the atlas library for \mb segmentation. LOOCV is conducted separately
for two different input datasets: the four Winterburn atlas subject BRAVO MR
images (referred to as the {\em WA BRAVO} dataset), and the five
Winterburn atlas subject images resampled to 0.9mm-isotropic voxel resolution
(referred to as the {\em WA Resampled} dataset. Each subject in the
dataset is segmented by \mb with that subject's image excluded from the atlas
library. The template library consists of 3T T1 images (0.9mm-isotropic
voxels) of healthy subjects in addition to the images from the dataset being
evaluated.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                               RESULTS                                        %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%                                                                     %%%%%
%%%%%     RESULTS: Experiment 1: Whole Hippocampus Cross-Validation       %%%%%
%%%%%                                                                     %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment 1 Results: Whole Hippocampus Cross-Validation}

@@

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: prep
% Prepares the data in a the form needed for plotting.
<<ADNI1-xval-prep,dependson='ADNI1-xval-load',include=FALSE,cache=TRUE>>=
all_data  <- read.csv('data/cache/ADNI-XVAL:all_data.csv')
all_data_mean  <- read.csv('data/cache/ADNI-XVAL:all_data_mean.csv')
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: mean Kappa plot by atlases, templates, reg and voting method. 
\begin{figure}
<<ADN1-xval-k-mean, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
ggplot(all_data_mean, 
       aes(x=templates.mb,y=k.mb, colour=as.factor(atlases))) + 
  facet_grid(reg_method~method.mb) + 
  stat_summary(fun.y=mean,geom='line',
               aes(y=k.mb, weight=1)) + 
  #stat_smooth(method='lm', formula=y~log(x)) + 
  scale_colour_hue(name="Atlases") +
  xlab( "Number of Templates" ) + 
  ylab( "Mean DSC" ) + 
  opts(legend.direction = "horizontal", legend.position = "bottom")
@
  \label{fig:ADNI1-xval-k_mean}
  \caption{
  {\bf Mean DSC of \mb segmentations across 69 ADNI1 subjects, by atlas and
  template library size, registration algorithm, and label fusion method.} 
  }
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: "increase" in mean kappa over multi-atlas
\begin{figure}
<<ADN1-Xval-k-diff,cache=TRUE,dependson='ADNI1-Xval-prep',fig.width=6,fig.height=4>>=
ggplot(subset(all_data_mean, reg_method=="ANTS" & method.mb=="Majority Vote" &
                templates.mb %in% seq(1,20,2)), 
       aes(x=templates.mb, y=k_diff, colour=as.factor(atlases))) + 
  #facet_grid(reg_method~method.y) + 
  stat_summary(fun.data=mean_cl_boot,geom='errorbar',
               aes(y=k_diff,colour=as.factor(atlases), width=0.2)) +
  stat_summary(fun.y=mean, geom="point", 
               aes(y=k_diff,colour=as.factor(atlases))) + 
  stat_summary(fun.y=mean, geom="line", 
               aes(y=k_diff,colour=as.factor(atlases))) + 
  #stat_smooth(method=lm, formula=y~log(x)) + 
  geom_hline(aes(alpha=0.5, yintercept=0), linetype='dashed') + 
  scale_x_continuous(breaks=seq(1,20,2)) + scale_y_continuous(breaks=seq(-1,1,by=0.01)) + 
  scale_colour_hue(name="Atlases") +
  scale_alpha_continuous(legend=FALSE) + 
  scale_linetype_discrete(legend=FALSE) + 
  xlab( "Number of Templates" ) + 
  ylab( "Difference in mean DSC" ) +
  opts(legend.direction = "horizontal", legend.position = "bottom")
@
  \caption{The difference in mean DSC between \mb and multi-atlas segmentations for 
  a range of parameter settings.}
  \label{ADNI1-xval-k-diff}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: show variability of MAGeT over parameters
\begin{figure}
<<ADN1-Xval-variability, cache=TRUE, dependson='ADNI1-Xval-prep'>>=
ANTSmajvote = subset(all_data, 
                     reg_method=="ANTS" & 
                     method.mb=="Majority Vote" &
                    templates.mb %in% seq(1,20,2))

@@

ggplot(stats.tests, 
  aes(x=templates.mb, y=p.value, colour=as.factor(atlases), 
      size = abs(statistic))) + 
  facet_grid( . ~ stat) + 
  geom_line(aes(size=0.1,area=0.1)) + 
  geom_point(aes(shape = factor(direction))) + 
  geom_hline(aes(yintercept=0.05, alpha=0.5), linetype='dashed') + 
  geom_hline(aes(yintercept=0.01, alpha=0.5), linetype='dashed') + 
  scale_shape_discrete("t-statistic sign", labels=c("positive","negative")) + 
  scale_size_area(name="t-statistic magnitude") + 
  scale_x_continuous(breaks=seq(1,20,by=2)) + 
  scale_y_log10(limits=c(0.001,1), breaks=c(0.01,0.05,0.1,1.0)) + 
  xlab( "Number of Templates" ) + 
  ylab( "p-value" ) +
  scale_colour_hue(name="Atlases") +
  scale_alpha_continuous(guide='none') + 
  scale_linetype_discrete(guide='none') + 
  scale_size_continuous(guide='none') + 
  opts(legend.direction = "horizontal", legend.position = "bottom")
@
  \caption{{\bf Difference in variability of \mb vs. multi-atlas segmentation 
  accuracy.}
  Variability of segmentation accuracy within 10 rounds of validation per subject. 
  Standard deviation and variance is computed per subject for both \mb and 
  multi-atlas, and the distribution of these test statistics is compared with a 
  t-test. The p-value of this test, is shown on the y-axis (scaled logarithmically).}
  \label{ADNI1-xval-variability}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-XVAL: Bland-Altman plots
\begin{figure}
<<ADN1-xval-Bland-Atlman, cache=TRUE, dependson='ADNI1-Xval-prep', fig.width=8, fig.height=8>>=
df_subset = subset(all_data, 
                      reg_method   == "ANTS" &
                      method.mb    == "Majority Vote" &
                      templates.mb %in% c(1,11,19) &
                      atlases      %in% c(1,5,9) &
                      volume.ma     < 10000)

@@

ggplot(hc_vol, aes(x=ma, y = diff, group=Atlases)) + 
    geom_hline(aes(yintercept = mean + c(-2, 0, 2) * sd, colour='brown'), 
                linetype=2) + 
    geom_point() + 
    geom_smooth(method="lm") + 
  facet_grid(Atlases~Templates, labeller = label_both) + 
  xlab(expression('Mean manual and MAGeT-Brain total hippocampus volume ' (mm^3))) + 
  ylab(expression('Manual - MAGeT-Brain total hippocampus volume ' (mm^3))) 
@
  \caption{Bland-Altman plots comparing manual and \mb hippocampal volumes 
  when using a varying number of atlases and templates.}
  \label{ADNI1-xval-Bland-Altman}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\FloatBarrier

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%                                                                     %%%%%
%%%%%    RESULTS: Experiment 2: Hippocampal Subfield Cross-Validation     %%%%%
%%%%%                                                                     %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment 2 Results: Winterburn Atlases Cross-Validation}
This experiment explores \mb segmentations of hippocampal subfields. To
achieve this, a leave-one-out validation is conducted in which lower-resolution
images ($0.9mm^3$ voxels) of each Winterburn atlas subject is segmented using 
the remaining Winterburn atlases. As a point of comparison, volumes of Winterburn 
atlases when downsampled to $0.9mm^3$ voxels are also computed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Winterburn-XVAL: compare resampling
\begin{figure}[h!]
<<WAval-vol-boxplot,cache=T,fig.width=6,fig.height=4>>=
data = subset(read.csv('data/WAval.csv'))
casted = dcast(melt(data, id.vars=c("file","version")), file + variable ~ version)
casted = rename(casted, c("variable"="region"))

levels(error$region) <-list("CA1"="X1", "CA1"="X101", 
                            "Subiculum"="X2","Subiculum"="X102", 
                            "CA4"="X4","CA4"="X104", 
                            "CA2/CA3"="X5","CA2/CA3"="X105", 
                            "Dentate"="X6","Dentate"="X106")
melted = melt(error,id.vars=c("file", "region"))
melted = rename(melted, c("variable"="measure", "value"="percenterr"))
levels(melted$measure) <- list("WA Resampled"="resampled",
                               "MAGeT + WA Resampled" = "mb_0.9mm",
                               "MAGeT + WA BRAVO"="mb_bravo")

ggplot(melted, aes(y=percenterr,x=region,colour=measure)) +
  geom_boxplot() +
  labs(y="Percent error in volume", x="Region") +
  scale_colour_discrete("") + 
  theme(legend.direction = "horizontal", legend.position = "bottom")
@
  \caption{{\bf Percent error in segmentation volume by hippocampus subregion.} 
  Percent error is measured against the volumes of of the unmodified Winterburn
  atlas segmentations.
  {\bf WA Resampled} are volumes of the manual segmentations of the Winterburn
  atlases after resampling to $0.9mm^3$.  
  {\bf MAGeT + WA Resampled} volumes are \mb segmentations of the Winterburn atlas
  images after resampling to $0.9mm^3$ voxels.
  {\bf MAGeT + WA BRAVO} volumes are \mb segmentations of T1 BRAVO images
  ($0.9mm^3$ voxels) acquired separately of four of the five Winterburn atlas
  subjects. 
  } 
  \label{fig:WAval-vol-boxplot}
\end{figure}

\FloatBarrier
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%                                                                     %%%%%
%%%%%    RESULTS: Experiment 3, Application of \mb to the segmentation    %%%%%
%%%%%                       of schizophrenia patients                     %%%%%
%%%%%                                                                     %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment 3 Results: Application to the segmentation first episode
\\ \mbox{schizophrenia patients}}

In this experiment \mb is applied to a dataset of images of first episode
schizophrenia parients, using the Winterburn atlases and a template library of
21 subject images selected at random.  Expert manual whole hippocampal
segmentations are used as gold standards.

\FloatBarrier
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%                                                                     %%%%%
%%%%%                           RESULTS                                   %%%%%
%%%%%                                                                     %%%%%
%%%%%         EXPERIMENT 3: Application of \mb to the segmentation        %%%%%
%%%%%                       of ADNI subjects                              %%%%%
%%%%%                                                                     %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Experiment 4 Results: Application to the segmentation of Alzheimer's
disease patients}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ADNI1-COMPLETE
% Prepares the data in a the form needed for plotting.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
<<ADNI-src-volumes-prep,echo=F,cache=T>>=
totals.complete  <- read.csv('data/cache/ADNI1:totals.complete.csv')
totals           <- read.csv('data/cache/ADNI1:totals.csv')
package_totals   <- read.csv('data/cache/ADNI1:package_totals.csv')
@
